### 2019-01-10 Thesis/thesis.tex
Made technical revisions.

Readers unfamiliar with game semantics should not worry too much about what $\bot$ is; the point of Proposition \ref{SamsonGuyAdequacy} is that the denotation of a command-type term precisely captures the operational behaviourof that term, in the sense that there is some distinguished strategy $\bot\from 1 \to \bC$ that is the denotation of every divergent program, and no convergent one.

@@

Working in the original category $\G$, however, we may identify this composite with the morphism
\[
  \deno{c\ts M[c/\sleep]} \from \bC \to \bC\,.
  \]
Now we have morphisms
\begin{mathpar}
  \deno{v\ts v \gets \suc !v} \from \Var \to \bC
  \and
  \deno{v\ts !v} \from \Var \to \bN\,,
\end{mathpar}
which together induce the composite
\[
  \Var \xrightarrow{\langle \deno{v\ts v \gets \suc !v},\deno{v\ts !v}\rangle}
  \bC \times \bN \xrightarrow{\deno{c \ts M[c/\sleep]} \times \id}
  \bC \times \bN \xrightarrow{\deno{\blank;\blank}}
  \bN\,.
  \]
Lastly, we may curry this term to give a morphism $1 \to (\Var\to \bN)$ and compose with the morphism
\[
  \new_\bN \from (\Var \to \bN) \to \bN\,.
  \]
The resulting morphism will be the denotation of the term
\[
  \new\;v=0\text{ in } M[v\gets \suc !v/\sleep];!v\from \nat\,.
  \]
Lemma \ref{IaSleepsSoundnessLemma} tells us that if $M\converges_n\skipp$ then this term converges to $n$ in Idealized Algol.

@@

\newcommand{\tit}{\mathrm{tt}}
\begin{definition}
  Let $\sigma\from 1 \to \bC$ be a Kleisli morphism in $\Kl_{R_{\bC}}$, considered as a morphism $\bC \to \bC$ in $\G$.  
  Define:
  \begin{IEEEeqnarray*}{rCl}
    \sigma' & = & \Var \xrightarrow{\langle \deno{v\ts v \gets \suc !v},\deno{v\ts !v}\rangle}
    \bC \times \bN \xrightarrow{\sigma\times\id}
    \bC \times \bN \xrightarrow{\deno{\blank;\blank}}
    \bN\,; \\
    \sigma'' & = & 1 \xrightarrow{\text{curry }\sigma''}
    (\Var \to \bN) \xrightarrow{\new_\bN}
    \bN\,.
  \end{IEEEeqnarray*}

  If there is some (necessarily unique) $n\in\omega$ such that
  \[
    1 \xrightarrow{\sigma''}
    \bN \xrightarrow{t_n}
    \bC
    \]
  is not equal to $\bot$, then we say that $n$ is the \emph{time taken for $\sigma$ to converge}, or $\tit(\sigma)$.
  If there is no such $n$, we say that $\tit(\sigma)=\infty$.
\end{definition}

@@

  Let us now work in the category $\G$.  
  So $\deno{M},\deno{N}$ may be regarded as morphisms $\bC \to \deno{T}$ and $\alpha$ may be regarded as a morphism $\deno{T} \to (\bC \to \bC)$, or equivalently as a morphism $\deno{T}\times \bC \to \bC$.
  Then the composites $\deno M;\alpha,\deno N;\alpha$ within $\Kl_{R_\bC}$ are given by the following composites in $\G$.
  \begin{mathpar}
    \bC \xrightarrow{\langle \deno M,\id\rangle}
    \deno T \times \bC \xrightarrow{\alpha}
    \bC
    \and
    \bC \xrightarrow{\langle \deno N,\id\rangle}
    \deno T \times \bC \xrightarrow{\alpha}
    \bC
  \end{mathpar} 
  By assumption, these composites are not equal in $\G$.  
  Then it is sufficient (since this is the case in $\G$) to assume that $\alpha$ is a compact strategy.
  Then, by the compact definability result in \cite{SamsonGuyIAActive}, $\alpha$ must be the denotation (in $\G$) of some term $P\from T \to \com \to \com$, which corresponds to the context
  \[
    C[x] = P\,x\;\sleep
    \]
  in $\Kl_{R_\bC}$.

  It follows that $\deno{C[M]} \ne \deno{C[N]}$.  
  Therefore, by Proposition \ref{IaSleepComputationalAdequacy}, $M$ and $N$ are not observationally equivalent.
\end{proof}

>>>>

To make sense of this strategy, we need to know that $\G$ is enriched in algebraic cpos.  
Then $\bot$ is the bottom element of $\G(1,\bC)$.

@@

Working in the original category $\G$, however, we may identify this composite with the morphism
\[
  \deno{\lambda c.M[c/\sleep]} \from 1 \to (\bC \to \bC)\,.
  \]
Now we have a morphism
\begin{mathpar}
  \eta = \deno{f\from \com\to \com \ts \lambda v.f(v\gets \suc!v);!v}\from (\bC \to \bC) \to (\Var \to \bN)\,.
\end{mathpar}
If we form the composite
\[
  1 \xrightarrow{\deno{\lambda c.M[c/\sleep]}}
  (\bC \to \bC) \xrightarrow{\eta}
  (\Var \to \bN) \xrightarrow{\new_\bN}
  \bN
  \]
then the morphism we get will be the denotation in $\G$ of the term
\[
  \ts \new\;v=0\text{ in } M[v\gets \suc !v/\sleep];!v\from \nat
  \]
(since the $\beta$ rule is valid for the semantics in $\G$).
Lemma \ref{IaSleepsSoundnessLemma} tells us that if $M\converges_n\skipp$ then this term converges to $n$ in Idealized Algol.

@@

\newcommand{\tit}{\mathrm{tt}}
\begin{definition}
  Let $\sigma\from 1 \to \bC$ be a Kleisli morphism in $\Kl_{R_{\bC}}$, considered as a morphism $1 \to (\bC \to \bC)$ in $\G$.

  If there is some (necessarily unique) $n\in\omega$ such that the composite
  \[
    1 \xrightarrow{\sigma}
    (\bC \to \bC) \xrightarrow{\eta}
    (\Var \to \bN) \xrightarrow{\new_\bN}
    \bN \xrightarrow{t_n}
    \bC
    \]
  is not equal to $\bot$, then we say that $n$ is the \emph{time taken for $\sigma$ to converge}, or $\tit(\sigma)$.
  If there is no such $n$, we say that $\tit(\sigma)=\infty$.
  \label{DefTimeTaken}
\end{definition}

@@

  Let us now work in the category $\G$.  
  So $\deno{M},\deno{N}$ may be regarded as morphisms $1 \to (\bC \to \deno{T})$ and $\alpha$ may be regarded as a morphism $\deno{T} \to (\bC \to \bC)$.
  Then the composites $\deno M;\alpha$ and $\deno N;\alpha$ within $\Kl_{R_\bC}$ are given by the following composites in $\G$.
  \begin{mathpar}
    1 \xrightarrow{\deno M}
    (\bC \to \deno T) \xrightarrow{\bC \to \alpha}
    (\bC \to (\bC \to \bC)) \xrightarrow{\mu}
    (\bC \to \bC)
    \and
    1 \xrightarrow{\deno N}
    (\bC \to \deno T) \xrightarrow{\bC \to \alpha}
    (\bC \to (\bC \to \bC)) \xrightarrow{\mu}
    (\bC \to \bC)
  \end{mathpar} 
  Without loss of generality, suppose that there exists some $n$ such that
  \begin{mathpar}
    \tit(\deno{M};(\bC \to \alpha);\mu)=n \and \tit(\deno{N};(\bC \to \alpha);\mu)\ne n\,.
  \end{mathpar}
  This means that we must have
  \begin{mathpar}
    \deno M;(\bC \to \alpha);\mu;\eta;\new_\bC;t_n \ne \bot
    \and
    \deno N;(\bC \to \alpha);\mu;\eta;\new_\bC;t_n = \bot
  \end{mathpar}
  in $\G$ (where $\eta$ is as defined in Definition \ref{DefTimeTaken}).

  $\G$ is enriched in algebraic cpos, so $\alpha$ is the least upper bound of its compact approximants.  
  It follows that there is some compact $\alpha'\subset \alpha$ such that
  \begin{mathpar}
    \deno M;(\bC \to \alpha');\mu;\eta;\new_\bC;t_n \ne \bot
    \and
    \deno N;(\bC \to \alpha');\mu;\eta;\new_\bC;t_n = \bot\,.
  \end{mathpar}
  In other words:
  \begin{mathpar}
    \tit(\deno M;(\bC \to \alpha');\mu) = n\,;
    \and
    \tit(\deno M;(\bC \to \alpha');\mu) \ne n\,.
  \end{mathpar}
  Now the compact definability result for Idealized Algol \cite[17]{SamsonGuyIAActive} tells us that $\alpha'$ must be the denotation of some IA term $x\from T \ts C[x] \from \com \to \com$, which is therefore the denotation of $C[\sleep] \from \com \to \com$ in $\Kl_{R_\bC}$.
  So we get
  \begin{mathpar}
    \tit(\deno{C[\sleep] M}) = n\,;
    \and
    \tit(\deno{C[\sleep] N}) \ne n\,.
  \end{mathpar}
  By Theorem \ref{IaSleepComputationalAdequacy}, $C[\sleep] M\converges_n$ and $C[\sleep] N\not\converges_n$.  
  Therefore, $M$ and $N$ are observationally inequivalent in IA+$\sleep$.
\end{proof}
